Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-13 Response.tex

Some additional detail has been included mentioning the difficulties in applying the \Poincare~section approach to the 3D case.

\begin{correction}
While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
\end{correction}

\item
    \begin{itshape}
-"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.  The drift in the state vector accuracy is still important to a mission designer, even if it is guaranteed to lie on an energy surface.  I think you should indicate that this is more for preliminary analysis to characterize the transfers, when this drift in state vector accuracy is tolerable.  I.E., it's useful for the introductory phases of mission design when trying to globally characterize the design space.  Characterizing the scope of the analysis in this way would be helpful to the audience.
\end{itshape}

The section on the variational integrator has been reduced and this specific sentence has been removed. 
However, additional detail has been provided in a different section of the paper which discusses this drawback of the variational integrator.

\begin{correction}
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
\end{correction}
\item 
    \begin{itshape}
"Optimal solutions are generally sensitive to small variations in the initial multipliers. As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance."  I don't agree that this implies that variational integrators are better for gradient-based methods.  The primary issue is that gradient-based methods that are solving optimal control problems are using first-order approximations to iterate on the problem.  The first-order approximation is what is poor, not necessarily the gradients.  Even if these gradients are of exact accuracy (and I agree that they should always be as accurate as possible), the first-order methods in the chaotic problem are what give the difficulties.  One could use a variational or standard integrator in a first-order gradient based method and nothing will change the fact that the method will be highly sensitive, save for partitioning the sensitivities as you did with your multiple shooting approach.  At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.
\end{itshape}

### 2017-07-13 manuscript.tex

>>>>

In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
